Mental health issue linked to major crime

### 2024-09-22 Paper/Paper.qmd
Add files via upload

---
title: "Mental health issue linked to major crime*"
author: Max LI
thanks: "Code and data are available at: https://github.com/ShaotongLi-Max/STA304-Paper-1-MaxLI."
date: 2024/9/21
date-format: long
abstract: "This Paper aims at discussing the relationship between Mental Health issue and committing crime. Using data from opendatatoronto, we can see Mental Health Act Apprehensions and major crime indicators of several police divisions in toronto. There appears to be positive linear relationship between the number of people having MHA apprehensions and number of crimes."
format: 
  pdf:
    toc: true           
    toc-title: "Table of Contents" 
    toc-depth: 2        
bibliography: references.bib
---

@@

ggplot(clean_data, aes(y = DIVISION, x = num_mha_reports)) +
  geom_bar(stat = "identity", fill = "#4E79A7", color = "black", width = 0.7) + # Custom color and bar width
  labs(title = "Number of MHA Reports by Division",
       y = "Police Division",
       x = "MHA Reports") +
  theme_minimal(base_size = 14) + # Improved theme with larger base font size
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), # Centered and bold title
    axis.text.y = element_text(size = 12), # Resized y-axis labels
    axis.title.y = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, max(clean_data$num_mha_reports) * 1.1)) + # Dynamic x-axis with padding
  geom_text(aes(label = num_mha_reports), hjust = -0.2, size = 4) # Display the values to the right of the bars

@@

ggplot(clean_data, aes(y = DIVISION, x = num_crimes)) +
  geom_bar(stat = "identity", fill = "#E15759", color = "black", width = 0.7) + 
  labs(title = "Number of Crimes by Division",
       y = "Division",
       x = "Crime Incidents") +
  theme_minimal(base_size = 14) + 
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), 
    axis.text.y = element_text(size = 12), 
    axis.title.y = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, max(clean_data$num_crimes) * 1.1), 
                     labels = abbreviate_number) + 
  geom_text(aes(label = abbreviate_number(num_crimes)), 
            hjust = -0.2, size = 4)

@@

```{r,fig.pos='H'}
#| label: Linear model summary
#| fig-cap: The summary table for linear model
#| echo: false
#| warning: false
#| message: false

@@

final_summary_df <- coefficients_df  # Only keep coefficients_df

summary_table <- gt(data = final_summary_df) %>%
  tab_header(title = "Linear Model Summary") %>%
  text_transform(
    locations = cells_body(columns = vars(term)),
    fn = function(x) ifelse(x == "(Intercept)", "Intercept", x)
  ) %>%
  fmt_number(
    columns = c(estimate, p.value),  # Only format the estimate and p.value
    decimals = 3
  ) %>%
  cols_label(
    estimate = "Estimate",
    p.value = "P-value"
  ) %>%
  cols_align(align = "center", columns = vars(estimate, p.value)) %>%
  tab_footnote(
    footnote = paste("Adjusted R-squared:", round(model_stats$adj.r.squared, 3)),
    locations = cells_title(groups = "title")
  ) %>%
  tab_footnote(
    footnote = paste("R-squared:", round(model_stats$r.squared, 3)),
    locations = cells_title(groups = "title")
  )

>>>>

---
title: "Mental health issue linked to major crime"
author: Max LI
thanks: "Code and data are available at: https://github.com/ShaotongLi-Max/STA304-Paper-1-MaxLI."
date: 2024/9/21
date-format: long
abstract: "This Paper aims at discussing the relationship between Mental Health issue and committing crime. Using data from opendatatoronto, we can see Mental Health Act Apprehensions and major crime indicators of several police divisions in toronto. There appears to be positive linear relationship between the number of people having MHA apprehensions and number of crimes."
format: 
  pdf:
    toc: true           
    toc-title: "Table of Contents" 
    toc-depth: 2        
bibliography: references.bib
---

@@

ggplot(clean_data, aes(y = DIVISION, x = num_mha_reports)) +
  geom_bar(stat = "identity", fill = "#4E79A7", color = "black", width = 0.7, position = position_dodge(width = 0.9)) +  
  labs(title = "Number of MHA Reports by Division",
       y = "Police Division",
       x = "MHA Reports") +
  theme_minimal(base_size = 14) + 
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5), 
    axis.text.y = element_text(size = 12),  
    axis.title.y = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_mha_reports) * 1.05)) +  # Adjust the padding on the x-axis
  geom_text(aes(label = num_mha_reports), hjust = -0.3, size = 3.5)

@@

ggplot(clean_data, aes(y = DIVISION, x = num_crimes)) +
  geom_bar(stat = "identity", fill = "#E15759", color = "black", width = 0.7) +  
  labs(title = "Number of Crimes by Division",
       y = "Division",
       x = "Crime Incidents") +
  theme_minimal(base_size = 14) + 
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),  
    axis.text.y = element_text(size = 12), 
    axis.title.y = element_text(face = "bold", size = 14),
    axis.title.x = element_text(face = "bold", size = 14)
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, max(clean_data$num_crimes) * 1.05), 
                     labels = abbreviate_number) +  
  geom_text(aes(label = abbreviate_number(num_crimes)), 
            hjust = -0.3, size = 3.5)

@@

```{r,fig.pos='H'}
#| label: Linear model summary
#| echo: false
#| warning: false
#| message: false

@@

final_summary_df <- coefficients_df

summary_table <- gt(data = final_summary_df) %>%
  tab_header(title = "Linear Model Summary") %>%
  text_transform(
    locations = cells_body(columns = vars(term)),
    fn = function(x) ifelse(x == "(Intercept)", "Intercept", x)
  ) %>%
  fmt_number(
    columns = c(estimate, p.value),  
    decimals = 3
  ) %>%
  cols_label(
    estimate = "Estimate",
    p.value = "P-value"
  ) %>%
  cols_align(align = "center", columns = vars(estimate, p.value)) %>%
  tab_footnote(
    footnote = paste("R-squared:", round(model_stats$r.squared, 3)),
    locations = cells_title(groups = "title")
  ) %>%
  tab_footnote(
    footnote = paste("Adjusted R-squared:", round(model_stats$adj.r.squared, 3)),
    locations = cells_title(groups = "title")
  ) %>%
  opt_footnote_marks(marks = "letters")
